### 2021-09-15 plos_latex_template.tex
sara's small comment on rule #10
Rocío: rewrote the introduction

%natalia: I changed the 1st sentence because we participated in the organizing team not just during the conference, but for several months before it. That was needed to work towards an inclusive conference

@@

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. The larger the conference, the larger the opportunities to network and learn from your cohort. However, conferences can become discriminating spaces, in which members of some specific privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities) reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.

While there has been some action to address systematic inequalities [where], there is still a lot of room for improvement.
% Rocío: Somehow this part below should be linked to the systemic inequalities part
Each one of us can have different points of view and experiences regarding inequality, depending on where we come from.
If you are reading this, you may be here because you already have some experience in Diversity, Inclusion, and Accessibility practices. Or you may want to improve diversity in your conference but do not know how or where to start. This article does not intend to be a comprehensive review of this topic --this would be impossible-- but will give some ideas pointing in key directions.   %Rocío: And finish with something more specific to connect with the next paragraph.

This article suggests rules to pivot traditional conferences towards inclusiveness and diversity. It is directed as people who are part of a stable meetings committee that oversees the site/location selection process or that coordinates with the local organizers (e.g. R Foundation Conference Committee, Ecological Society of America Meetings Committee). It is also directed at the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.

These tips stem from the authors' experience of organizing useR! 2021, a virtual and global conference for users and developers of the R programming language \cite{r_core_team_2021}. We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone. Most of the authors also have experience of organizing other regional and national academic conferences and events in communities of practice such as R-Ladies and The Carpentries. Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.

\section{Define what diversity and inclusion mean for your conference}
\label{rule_diversity}

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. Human diversity should be celebrated and respected in every way. Nonetheless, there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built. While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

Depending on the field, scientific or professional communities can have very different compositions. Thinking about which subgroups have been historically underrepresented in your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event. Doing this exercise while you ...
Think about the reasons why you want to create a more diverse conference can avoid falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes., and do not act as a benefactor or savior but as a genuine collaborator.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference. Privilege is largely invisible to those who have it. Once we understand our own privileges, we can begin addressing it on an individual and institutional basis.
%Natalia: I'll look for a reference for privilege definition.
%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.  I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule. Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."

Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context. Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision. Would this translate into an even gender distribution in your speakers? Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees? Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions? This vision should guide and help hold the organizing team accountable along the way.

@@

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTIQA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real. Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

Paying attention to some dimensions of diversity while neglecting others may have unintended consequences. Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. In severe cases, they can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.

Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of  attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, this sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community. Some examples are the newbies sessions for first timers, mixers lead by R-Ladies or MiR, or a leisure activity that reunites subgroups with the same interests: arts, exercise, sports, movies, etc.

@@

Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups. First, delegate. Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer coordinating positions, even if this means stepping down. Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others. Break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while men take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).

Words matter. In addition to excluding derogatory or discriminatory language (remember that the code of conduct applies to you too, Rule \ref{rule_inclusion}), make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept any feedback without being defensive.

@@

%Rocío: Yani's ref to have just in case:
% https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters
%Natalia: I don't know if this is the best place, but I'd add the term "well-being".

@@

In-person interaction is priceless; however, it is more expensive for some, and even unattainable for others. This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities. This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}. Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.

@@

Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles, and accessibility needs. Social events in presencial conferences are often the most exclusive, and they are a very common place where you have code of conduct violations. Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond. An online event is a great opportunity to create spaces for people who do not enjoy in person networking. Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.

@@

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.  An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference.Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.

@@

Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them. If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g. `grant recipient' or `diversity scholar', respectively). Do not call it a scholarship or a grant. Ultimately, it is a fee waiver or a discount for the people who need additional financial support. Also, the process for applicants should be simplified. Applying for loans, grants, and scholarships may be an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility. If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual). If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions or design accessible presentations.

Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panellist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

\section{Actively reach out and promote the conference to people who have been systematically excluded}
\label{rule_communication}
%Andrea: this rule name should be more explicit towards communication, "promoting"
If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines. Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory against underrepresented groups. Advertise the conference in multiple languages (rule \ref{rule_language}). Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too. 
% Yani: We need to work a little more here, perhaps with some examples.
% Rocío: could add some ideas from here 
% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf

\section{Do not expect the organizing team to work full time and for free}
\label{rule_pay_team}

Having a diverse team and executing inclusive and accessible practices throughout the organizing period --it might be a year or two --may require a lot of effort. The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. However, having a strong community as the only reward may be enough if you are in a privileged position. Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. In addition, tasks such as taking and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded. Be mindful of each particular context, be flexible with hours or commitments, and revise your budget; if possible, secure funds to offer stipends to them early on.

>>>>

@@

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. 
However, conferences are likely to reproduce the systematic discrimination occurring in other spaces in our fields. %RJ: (Biggs et al. 2018) https://link.springer.com/article/10.1007/s11199-017-0800-9
Conversely, efforts to address systematic discrimination in conferences and make them inclusive spaces can have a meaningful impact in our communities and the lives and careers of their members.

This article suggests rules to pivot traditional conferences towards diversity and inclusion, and strive to build more inclusive and welcoming communities. 
Some scholarly or technological communities are formally structured as learned societies (e.g. the International Geographical Union or the Royal Statistical Society), others are composed by networks of less formal local groups (e.g. R User Groups or Python User Groups), and others may not have any established structure. 
For the latter, conferences may play even a bigger role in shaping the community, since there is no other organization or institution to gather the community members.

The rules written here are directed to people who are part of a stable meetings committee that oversees the site/location selection process, that coordinates with the local organizers of conferences, or even the local/virtual organizers who desire to make an inclusive conference starting at the planning stage.
These tips stem from the authors' experience of organizing useR! 2021, a virtual and global statistical computing conference for users and developers of the R programming language \cite{r_core_team_2021}. 
We embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, truly inclusive and accessible experience for everyone. 
Here, we share the lessons learned within the past year of organizing this global useR!2021, summarized as 10 simple rules towards an inclusive conference.

%   However, conferences can become discriminating spaces, in which members of some specific privileged groups reproduce the systematic inequalities that occur in academia and society \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.

\section{Embrace all dimensions of diversity}
\label{rule_diversity}
%ast at the end i think we should go back to Embrace all dimensions of diversity
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. Human diversity should be celebrated and respected in every way. Nonetheless, there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems --including conferences-- are consciously and unconsciously built. While no isolated initiative can change reality by itself, building a more diverse and inclusive conference starts by recognizing that these inequalities have excluded systematically whole groups of people from academia and from scientific and professional circles \cite{timperleyHeMoanaPukepuke2020}.

Depending on the field, scientific or professional communities can have very different compositions. Thinking about which subgroups have been historically excluded from your community, and thinking about which subgroups hold most privileges can help you guide the organization of a more inclusive and diverse event. Doing this exercise while you think about the reasons why you want to create a more diverse conference can prevent falling into the mistake of treating diversity as a checklist, or using diversity and inclusion work for image purposes.
%andrea: this is here because we want to avoid readers to read as recipes and treat this less seriously than how it should (and then blaming the recipe). i am not sure of the phrasing or position in the text."genuine collaborator" is in the spirit of co-liberation/co-conspirator
%Yanina: They are going to blame what they want to blame. You can't control what people are going to do and say and we shouldn't write defensively because we can't take up space in our message with that. People who do not want to change will not change because of our paper. I imagine this as a guide and help for those people who do want to do something and want to do their best and do not know what or how to do it. Or they have already done something and want to continue deepening those changes. The rest is not worth spending time and energy.  But perhaps the goals of the paper change in some other discussion and I lost that.
Actively be an ally: if you understand your own privileges (unearned advantages given by society to some people but not all) \cite{}, you can help to work on end oppression in your conference. Privilege is largely invisible to those who have it. Once we understand our own privileges, we can begin addressing them on an individual and institutional basis.
%Natalia: I'll look for a reference for privilege definition.
%marcela: I modified the paragraph a bit, but it needs more work to connect Natalia's and Andrea's ideas. It's just a suggestion, feel free to re-write it again.
%Batool: I tried making some paraphrasing - It's just a simple suggestion, not sure if it deliver the message you mean it.  I also think that we may need to define "Privilege" at the 1st mention in the 1st paragraph in this rule. Something like "Privilege is a set of unearned benefits enjoyed by a group, beyond what’s available to others."

Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context. Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your vision. Would this translate into an even gender distribution in your speakers? Would it be representation of all races -especially Black people-- among the head organizers, speakers, and attendees? Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions? This is not an invitation to prioritize some dimensions of diversity above others This vision should guide and help hold the organizing team accountable along the way.

@@

%Yanina: It's probably a language problem but how I read this rule It is like to give the organizers a carte blanche to define what diversity and inclusion is and with that they are covered. What if they decide that everything is fine if they are all white people, native English speakers and without disabilities? I don't think we want to encourage that.  I mean, you need a bottom line and from them you build to improve. Plus, as it is well highligth in these rules, it is difficult to realize our own privileges, so, again, you cannot have a definition of diversity and inclusion for the conference if the team that defines it is not diverse. How are you going to have different ideas from the ones you generally have?

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable are just some examples of decisions that can make inclusion real. Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.

Nevertheless, having a diverse offer of networking opportunities that can appeal to people with different backgrounds, accessibility needs, and preferences can be challenging, especially if you have the mindset of organizing each activity for the complete pool of attendants. If you think of smaller activities that can reach specific groups, preferably co-lead by community leaders, these sessions can be more productive and successful, and even inclusive, than trying to organize one single activity that can please the whole community. Some examples are the newbies sessions for first-timers, mixers lead by specific subgroups or communities, or leisure activities that reunite subgroups with the same interests: arts, exercise, sports, movies, etc.

@@

Most advice when working with a diverse team should apply to every team, but is especially true when working with people from marginalized groups. First, delegate. Don't expect self-nomination and voting to work as mechanisms to counteract systemic inequalities. Nominate directly and offer coordinating positions, even if this means stepping down. Offer mentorship and guidance if you encounter cases of impostor syndrome, doubts about the use of English in communication (See Rule \ref{rule_language}), or others. You can, for example, break the expectations about leadership as a lonely task and create smaller, co-led groups, where everyone finds their preferred tasks and gets to take leadership. Splitting the workload and responsibilities should not be done by putting care-taking labors --community building, meeting organization and note taking, conversations with potential partners-- on the hands of women and other minoritized groups, while people from privileged groups take the lead in stereotypical highly-valued tasks (see Rule \ref{rule_spotlight}).

Set the tone. %adding here from Liz's rule but maybe this could go in communication in general
In addition to excluding derogatory or discriminatory language (remember that the code of conduct applies to you too, Rule \ref{rule_inclusion}), make the effort to teach yourself the vocabulary and the best ways to communicate to account for every culture and situation. Do not expect people to teach you --it's not their role-- and accept feedback without being defensive.

@@

%Rocío: Yani's ref to have just in case:
% https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters

Make sure that every selection committee --the committee looking for keynote speakers, the selection committee for abstracts, the prizes and award committees --are also diverse \cite{swartzScienceValueDiversity2019, wongBuildDiversityScience2020}, and ask them to be aware that everyone has an implicit biases, to recognize them, and try to counteract them.  An inclusive and diverse organizing team (\textbf{Rule \ref{rule_organizing_team}}) is already a great starting point to overcome this bias in other roles. The regional and local communities in your field are also good sources to tap into. 
For example, for useR! 2021, groups like AfricaR (africa-r.org), R-Ladies (rladies.org), MiR (mircommunity.com), Forwards (forwards.github.io), and LatinR (latin-r.com) were fundamental to reach people for the organizing team, potential presenters, to co-lead social events, and attendees; some of them even held spaces to help the members of their groups to prepare abstracts for submission.
When organizing an international academic conference, you may reach local associations, student organizations or other Early Career Researchers groups or communities of practice: ask their needs and barriers, and explicitly invite them to contribute to the conference. Since many conferences are held with the primary goal of displaying cutting-edge work led by senior scientists and academics, early career researchers are often disadvantaged. Diversity should spread among all speaking roles, and the conference should also help raise the profiles of early career researchers, fostering collaboration and building their skills.

@@

In-person interaction is priceless; however, going to conferences is expensive, and many people do not attend for lack of financial support. This inaccessibility %ast: maybe inaccessibility is not the word here because we want to focus on disabilities with this word
is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. Online conferences are more inclusive: they do not require a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities. This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}. \cite{https://link.springer.com/content/pdf/10.1007/s10869-011-9232-7.pdf} found that the order of communication medium influenced perceived inclusion, which in turn influenced individual participation.  The study suggests that within diverse groups in which minority member is different from the rest of team and may feel excluded, initial computer-mediated communication may help the minority member feel more included. Furthermore, an online format is more environmentally-friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts2021}.

%Yanina: I add the reference about people feeling better using on-line and computer mediated communication when belong to a excluded group.

@@

Take advantage of the online setting for new networking activities that can be inclusive for people with different socialization styles and accessibility needs. Social events in in-person conferences are often the most exclusive, and they are a very common place where you have code of conduct violations. Having them directly translated into their virtual versions does not address the need for creating opportunities for people to meet and bond. An online event is a great opportunity to create spaces for people who do not enjoy in person networking. Some ideas in this line are: offering the option of written chat only, instead of voice or video conversations, opening events with team work like trivia, offer some events that can be passively enjoyed like movies, yoga sessions, or art displays, where attendants can choose to just sit and enjoy without talking, or have a chat channel to comment on their experiences during the session. Respect people’s limits, preferences, and remember that “the usual” does not necessarily work with your community, and that one single networking activity will never serve the whole community. It is worth trying new and varied activities that might work with subgroups of attendees.

@@

Meeting the need of linguistically diverse audience is an important aspect in inclusion that is often overlooked. Non-native English speakers may feel intimidated to ask questions or raise their views. Providing a welcoming and diverse environment by encouraging the full participation of those individuals is critical (Rule 2). If the panelist or speakers will not be presenting in their native language, it may helpful to rephrase audience questions in a less complex language. Consider letting speakers know in advance that they can ask for a question to be translated to them in their native language and that they can respond in their native language.

\section{Express this welcoming spirit in your communication strategy}
\label{rule_communication}
%Andrea: this rule name should be more explicit towards communication, "promoting"
\begin{itemize}
    \item Promotion: Actively reach out and promote the conference to people who have been systematically excluded/Outreach/promotion plan that reaches communities that are traditionally excluded in scientific meetings
    \item Express this welcoming spirit in your communication strategy/Be public about your commitment to equity and inclusion: display the coc prominently in conference spaces, publish the diversity statement, fee waivers, abstract selection process
    \item Languages
    \item Words matter: Be careful how you talk about minoritized people, Assertively set tone of inclusion
\end{itemize}

If part of the community has been historically discriminated against, one should emphasize that they are particularly welcome in this event and that the organization will make it a safe and inclusive environment. Publish the diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}), the code of conduct and accessibility guidelines. Fee waivers, financial support, and ease of application/registration should also be advertised; people who fear getting rejected when applying for scholarships and waivers may find it relieving that the process will be supportive rather than discriminatory. Advertise the conference in multiple languages (rule \ref{rule_language}). Express this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) to let people know that they are seen, respected, and welcome; that this is their space and community too. 
% Yani: We need to work a little more here, perhaps with some examples.
% Rocío: could add some ideas from here 
% https://www.science.org.au/files/userfiles/support/emcr/documents/one-page-summary-emcr-improving-diversity-web.pdf

@@

Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them. If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or terms that single out the recipients (e.g. `grant recipient' or `diversity scholar', respectively). Do not call it a scholarship or a grant. Ultimately, it is a fee waiver or a discount for the people who need additional financial support. Also, the process for applicants should be simplified. Applying for loans, grants, and scholarships may be an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility. If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees and visa financial and administrative support (if in-person), or internet connection services (if virtual). If the conference is virtual, you may offer technical support so that attendees can record their videos, include captions, or design accessible presentations.

\subsection{Do not expect the organizing team to work full time and for free}
\label{rule_pay_team}
Having a diverse team and executing inclusive and accessible practices throughout the organizing period --it might be a year or two --may require a lot of effort. The effort is worth it because it strengthens the event and the community making it truly welcoming for everyone. However, having a strong community as the only reward may be enough if you are in a privileged position. Some people and often the minoritized ones do not have institutional support to put time and effort into the organization tasks and do not have the luxury to commit to the organization for free. In addition, tasks such as taking and responding to code of conduct reports, can be emotionally intense work and should be additionally rewarded. Be mindful of each particular context, be flexible with hours or commitments, and revise your budget; if possible, secure funds to offer stipends to them early on.

\section{Diversity and inclusion are processes}
\label{rule_process}

%Sara: I think the last rule should be about the process and reinforce to actually discuss and include diversity as a principle and not as a checklist (as it is on the text). Maybe also something about accountability and permanence. I think that this rule go be together with the rule_organizing_team given that it is relative to the team work and their dynamics.
%Sara: copying Yani comment here -  I like the "accountable and permanent" name. I think that an idea about progressive improvement can be mention here, like improve something every edition and no go back on the conquers goals. Now, to be accountable, you need to be transparent, and you need a clear structure; you get that with the structure of your team, with your declaration and guidelines, and with your action to reinforce those declarations and guidelines (if you have only declarations then you have only empty words, that sound good but you no need to comply. We discuss a lot about this).  For me, one more time, putting together a diverse team is what makes the other thing happened. We can cite The tyranny of structurelessness. The Second Wave, 2(1), 1972. on this rule. Perhaps we can talk about governance.
%ast: i'll put the last version of the tenth rule just to have an idea of how we left this
%past names: Have a bold vision but be ready to compromise when necessary/Aim for progress not perfection
%more people rules that have a continuity/accountability theme
%Aim for Progress, not Perfection
%Plan, balance and keep working to improve inclusion in the next edition of the conference!
%Assess whether equity and inclusion goals were met during the meeting
%Be accountable and permanent
%Use feedback from previous conferences or other conferences in your field to identify barriers to inclusion and plan around them

This rule is a complement to all the other rules. It is important to have a bold vision and clear targets. Some parts might be defined as baseline assumptions that are not up for discussion. However, there will always be compromises. Ideal solutions that meet all our expectations are often not possible. Don't let the perfect be the enemy of good. As a team, this is a challenging journey. Very early on, we tried to recognize that our resources were limited (both in time of the organizing team but also in money). We had to let go of certain aspects that were important to many of us. For an inclusive conference, we need to avoid giving in because considering all aspects seems impossible and burning ourselves and the team out because considering all aspects is actually impossible.
